Enhancing the prediction of sales for newly published books by means of computational intelligence methods

### 2016-04-23 paper.tex
new description for SOM

Kohonen's SOM is a feedforward neural network \cite{Haykin98_NNC} that
uses an unsupervised training algorithm and which, through a process
called self-organisation, configures the output units into a
topological representation of the original data. It tries to imitate
the self-organisation done in the sensory cortex of the human brain,
where neighbouring neurons are activated by similar stimulus. SOM
belongs to a general class of neural network methods, which are
non-linear regression techniques that can be trained to learn or find
relationships between inputs and outputs or to organise data so as to
disclose so far unknown patterns or structures. It is usually used as
a clustering/classification tool or used to find unknown relationships
between a set of variables that describe a problem. The main property
of the SOM is that it makes a nonlinear projection from a
high-dimensional data space (one dimension per variable) on a regular,
low-dimensional (usually 2D) grid of neurons, called units. 
SOM is further processed using Ultsch method \cite{UmatUlts}, the
Unified distance matrix (U-matrix), which uses SOM's codevectors
(vectors of variables of the problem) as data source and generates a
matrix where each component is a distance measure between two adjacent
neurons. It allows us to visualise any multi-variated dataset in a
two-dimensional display, so we can detect topological relations among
neurons and infer about the input data structure. High values in the
U-matrix represent a frontier region between clusters, and low values
represent a high degree of similarities among neurons on that region,
clusters .

SOMs (and U-Matrix) are usually applied to visualise natural
structures in the data and their relations, as well as the natural
groupings that could be among them. In addition, SOM makes easy the
estimation of the variables that have more influence on these
groupings. Other statistical and soft computing tools can also be used
for this purpose, but since Kohonen's SOMs offers a visual way of
doing it, it is much more intuitive, and takes advantage of the
capabilities of the human brain as a pattern recogniser. 
%FERGUSON: por qué se menciona aquí la palabra "Kohonen" y no antes?
% Antonio - arreglado 
% And this is going to be useful because... - JJ

>>>>

%--- new SOM description % -Sam

SOM is a type of unsupervised neural networks inspired by the topographical organization of the sensory cortex of human brain \cite{kohonen1998}. It maps highly dimensional data into a lower dimensional space where similar data patterns are located in more adjacent areas on the grid.

SOM is a very useful analysis tool for visualizing complex data with high number of variables in a simple geometric planes. Therefore, it could be used for different tasks like features reduction and data clustering and visualization.

Basically, SOM consists of two layers; input layer and output layer (also known as Kohen layer). Input layer is fully connected with the Kohen layer which is usually a two dimensional grid of neurons (i.e.; units). This means that every neuron is linked to every input in the input layer.

The learning process of SOM is an iterative process applied in two stages. First, the weights of the network are initialized randomly then the n-dimensional input vectors are presented to the network sequentially so the distance between the weights of the neurons and the inputs are computed using the simple Euclidean distance. In the second stage, the neuron which has the shortest distance with the presented input vector is updated along with its neighbor neurons. This neuron is called the best matching unit (BMU). The goal of the learning process is to minimize the neighborhood distance in the output clusters.

When the learning process is done a visual representation method like the Unified Distance Matrix, or U-matrix for short, is used to visualize the distances between neurons in the output space. In the U-matrix, the distance between adjacent neurons is measured and assigned different colors. Dark colors between neurons indicate large distances (boundaries) while light colors represent similar areas (clusters).

%---

%FERGUSON: por qué se menciona aquí la palabra "Kohonen" y no antes?
% Antonio - arreglado 
% And this is going to be useful because... - JJ
